---
title: "Panel Reactive Antibodies (PRAs) and Their Association with Blood Product Exposures in Pediatric Patients Undergoing Cardiac Transplantation"
author:
  - 'Dr. Nina Guzzetta, Dr. Susana Cruz Beltran'
  - 'Statisticians: Dr. Renee Moore, Andrea Lane'
date: "3/10/2021"
geometry: margin=1cm
output:
  pdf_document:
    toc: true
    number_sections: true
header-includes:
    - \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
library(readxl)
library(epitrix)
library(dplyr)
library(rockchalk)
library(kableExtra)
library(summarytools)
library(ggplot2)
library(ggmosaic)
library(tidyverse)
library(berryFunctions)
library(caret)
library(glmnet)

options(knitr.table.format = "latex")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")

```

\captionsetup[table]{labelformat=empty}

```{r upload}
df.sheet1 <- read_excel("Transplant data.final.xlsx", sheet = 1,
                        col_types = c("text","text","date","text","numeric",
                                      "text","text","text", "date","date",
                                      "text","text","text","numeric","text",
                                      "text",rep("numeric",5),"text",rep("numeric",16)))
df.sheet2 <- read_excel("Transplant data.final.xlsx", sheet = 2,
                        col_types = c("text","text","date","text","numeric",
                                      "text","text","text", "date","date",
                                      "text","text","text","numeric",rep("numeric",6),"text",
                                      "text",rep("numeric",5),"text",rep("numeric",16)))

df.neg <- df.sheet1[-1,]
df.pos <- df.sheet2[-1,]

names(df.neg) <- clean_labels(names(df.neg))
names(df.pos) <- clean_labels(names(df.pos))

names(df.neg)[1] <- "Subj_num"
names(df.pos)[1] <- "Subj_num"

names(df.neg)[25] <- "prior_rbc"
names(df.neg)[26] <- "prior_ffp"
names(df.neg)[27] <- "prior_plt"
names(df.neg)[28] <- "prior_cryo"

names(df.neg)[30] <- "dos_rbc"
names(df.neg)[31] <- "dos_ffp"
names(df.neg)[32] <- "dos_plt"
names(df.neg)[33] <- "dos_cryo"

names(df.neg)[35] <- "post_rbc"
names(df.neg)[36] <- "post_ffp"
names(df.neg)[37] <- "post_plt"
names(df.neg)[38] <- "post_cryo"


names(df.pos)[31] <- "prior_rbc"
names(df.pos)[32] <- "prior_ffp"
names(df.pos)[33] <- "prior_plt"
names(df.pos)[34] <- "prior_cryo"

names(df.pos)[36] <- "dos_rbc"
names(df.pos)[37] <- "dos_ffp"
names(df.pos)[38] <- "dos_plt"
names(df.pos)[39] <- "dos_cryo"

names(df.pos)[41] <- "post_rbc"
names(df.pos)[42] <- "post_ffp"
names(df.pos)[43] <- "post_plt"
names(df.pos)[44] <- "post_cryo"

df.all <- bind_rows("1" = df.neg, "2" = df.pos, .id="pra_status")

```

```{r clean}
df.all$age_unit <- substr(df.all$age_years,nchar(df.all$age_years),nchar(df.all$age_years))
df.all$age_num <- substr(df.all$age_years,1,nchar(df.all$age_years)-1)

df.all$age <- ifelse(df.all$age_unit == "m", as.numeric(df.all$age_num)/12, ifelse(df.all$age_unit == "w", as.numeric(df.all$age_num)/52, as.numeric(df.all$age_num)))

df.all$abo_incompatible_nomiss <- ifelse(is.na(df.all$abo_incompatible),0,1)
df.all$prior_mech_nomiss <- ifelse(is.na(df.all$prior_mech_support),0,1)

df.all$plasma_c <- ifelse(df.all$plasmapheresis %in% c("Yes","yes"),1,0)
df.all$rej_c <- ifelse(df.all$rejection=="N",0,1)

df.all$plasma_c <- factor(df.all$plasma_c, levels = c(0,1),
                          labels = c("No","Yes"))
df.all$rej_c <- factor(df.all$rej_c, levels = c(0,1), labels = c("No","Yes"))
df.all$ecmo_c <- factor(df.all$ecmo_support_postop, levels = c("no","yes"),
                        labels = c("no","yes"))
df.all$diag_c <- df.all$diagnosis
df.all$diag_c[df.all$diag_c=="LV failure"] <- "Dilated CM"
df.all$diag_c[df.all$diag_c=="RV Failure" | df.all$diag_c=="Shone's complex- Coarct - Multiple VSD"] <- "Other"

df.all$diag_c <- factor(df.all$diag_c, 
      levels = c("ARVD","Dilated CM","Failed SV","Restrictive CM","Other"),
      labels = c("ARVD","Dilated CM","Failed SV","Restrictive CM","Other"))
df.all$donor_c <- factor(df.all$donor_specific_ab, levels = c("N","Y"),
                         labels = c("No","Yes"))
df.all$gender_c <- factor(df.all$gender, levels = c("F","M"), labels = c("Female","Male"))
df.all$race_c <- factor(df.all$race, 
        levels = c("White","African American","Hispanic","Asian","Declined"),             labels = c("White","African American","Hispanic","Asian","No response"))
df.all$abo_incompatible_nomiss <- factor(df.all$abo_incompatible_nomiss, 
                                         levels = c(0,1), labels = c("No","Yes"))
df.all$prior_mech_nomiss <- factor(df.all$prior_mech_nomiss, levels = c(0,1),
                                         labels = c("No","Yes"))

df.all <- df.all[!is.na(df.all$prior_donor_exposures),]
```

```{r functions}
my2x2 <- function(X,Y,col=TRUE){
  perc <- ifelse(col,2,1)
  mytab <- table(X,Y)
  proptab <- format(round(prop.table(mytab,perc)*100,1),nsmall = 1,trim=TRUE)
  mytab2 <- matrix(paste0(mytab," (",proptab,")"),nlevels(X),2)
  mytab2 <- rbind(mytab2, colSums(mytab))
  mytab <- cbind(mytab2, c(rowSums(mytab),sum(rowSums(mytab))))
  return(mytab)
}

mypval <- function(X,Y,symbol=TRUE){
  if (class(X)=="numeric"){
    test <- "Wilcoxon rank-sum"
    pval <- wilcox.test(X[df.all$pra_status==1],X[df.all$pra_status==2])$p.value
    pval <- ifelse(pval == 1,0.99,pval)
    pval <- format(round(pval,2), nsmall=2)
    pval <- ifelse(pval==0,"<0.0001",pval)
  } else if (class(X)=="factor"){
  
  tab <- table(X,Y)
  exp.tab <- chisq.test(tab)$expected
  test <- ifelse(all(exp.tab > 5),"Chi Square","Fisher's Exact")
  pval <- ifelse(all(exp.tab > 5),chisq.test(tab,correct = FALSE)$p.value,fisher.test(tab)$p.value)
  pval <- ifelse(pval == 1,0.99,pval)
  pval <- format(round(pval,2), nsmall=2)
  pval <- ifelse(pval=="0.00","<0.01",pval)
  if(symbol){pval <- ifelse(test=="Chi Square",
                 paste0(pval,""),
                 paste0(pval,"*"))}
   }
  return(list(pval = pval, test = test))
}

varlevels <- function(X){
  if(class(X)=="numeric"){
    nlev <- 1
  } else if (class(X)=="factor"){
    nlev <- nlevels(X)
  }
  return(nlev)
}

mystats <- function(X,Y){
  if (class(X)=="numeric"){
    tab <- matrix(NA,1,3)
    all_1 <- format(round(median(X,na.rm = TRUE),1),nsmall = 1)
    all_2 <- paste0("(",format(round(quantile(X,na.rm = TRUE)[2],1),nsmall=1),", ",
                    format(round(quantile(X,na.rm = TRUE)[4],1),nsmall=1),")")
    
    strat1_1 <- format(round(median(X[Y==1],na.rm = TRUE),1),nsmall=1)
    strat2_1 <- format(round(median(X[Y==2],na.rm = TRUE),1),nsmall=1)
    
    strat1_2 <- paste0("(",format(round(quantile(X[Y==1],na.rm = TRUE)[2],1),nsmall=1),", ",
                       format(round(quantile(X[Y==1],na.rm = TRUE)[4],1),nsmall=1),")")
    strat2_2 <- paste0("(",format(round(quantile(X[Y==2],na.rm = TRUE)[2],1),nsmall = 1),", ",
                       format(round(quantile(X[Y==2],na.rm = TRUE)[4],1),nsmall = 1),")")
    
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)
    
    tab <- cbind(cell1,cell2,cell3)
  } else if (class(X)=="factor"){
    nlev <- nlevels(X)
    all_1 <- c()
    all_2 <- c()
    strat1_1 <- c()
    strat2_1 <- c()
    strat1_2 <- c()
    strat2_2 <- c()
    tab <- matrix(NA,nlev,3)
    for(i in 1:nlev){
    tab1 <- table(X)
    all_1[i] <- paste(tab1[i])
    all_2[i] <- paste0("(",format(round(prop.table(tab1)[i]*100,1),nsmall = 1),")")
    
    tab2 <- table(X,Y)
    strat1_1[i] <- tab2[i,1]
    strat1_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,1]*100,1),nsmall = 1),")")
    
    strat2_1[i] <- tab2[i,2]
    strat2_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,2]*100,1),nsmall = 1),")")
      
    }
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)

  }
  return(tab)
}

```

\clearpage
\section{Tables}
```{r table1}
varlist <- c("age","weight","no_of_prior_surgeries","gender_c","race_c","prior_mech_nomiss","diag_c","donor_time","cpb_time","abo_incompatible_nomiss","plasma_c","icu_length_of_stay_day", "vis", "time_to_extubation_hr","rej_c", "donor_c","ecmo_c")
# var_levels <- lapply(df.all[varlist], FUN = varlevels)
# df_levels <- data.frame(matrix(unlist(var_levels), nrow=length(varlist),byrow = T))

myfulltab <- lapply(df.all[varlist], FUN = mystats, Y = df.all$pra_status)

mat.tab <- as.matrix(unlist(myfulltab[[1]]))
for (i in 2:length(myfulltab)){
  mat.tab <- rbind(mat.tab,as.matrix(unlist(myfulltab[[i]])))
}

mypvals <- lapply(df.all[varlist], function(x)
                  rep(mypval(x,df.all$pra_status)$pval,varlevels(x)))
# mytest <- lapply(df.all[varlist], function(x)
#                   rep(mypval(x,df.all$pra_status)$test,varlevels(x)))
# mytestmat <- as.matrix(unlist(mytest))
mypvalmat <- as.matrix(unlist(mypvals))
mat.tab.p <- cbind(mat.tab,mypvalmat)
df.tab <- data.frame(mat.tab.p)

names(df.tab)[2] <- paste0(names(df.tab)[2],
  footnote_marker_symbol(1, "latex"))
rownames(df.tab) <- c("Age (Years)", "Weight at transplantation (kg)", 
                      "Number of previous surgeries","Female", "Male", "White",
                      "African American", "Hispanic","Asian", "No response",  
                       "No", "Yes","ARVD","Dilated CM","Failed SV",
                      "Restrictive CM","Other","Donor cross clamp time (min)",
                      "CPB time (min)","No ","Yes ","No  ","Yes  ",
                      "Length of ICU stay (days)","Vasoactive inotrope score",
                      "Time to extubation (hrs)","No   ","Yes   ","No    ",
                      "Yes    ","No     ","Yes     ")

kable(df.tab[1:17,], "latex", booktabs = T, align = "l", col.names = c("All patients (n=64)", "PRA- (n=35)", "PRA+ (n=29)", "p value"), caption = "Table 1a: Demographic and Preoperative Variables") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Gender", 4, 5, bold = FALSE) %>% 
  group_rows("Race", 6, 10, bold = FALSE) %>% 
  group_rows("Mechanical support prior to transplantation", 11, 12, bold = FALSE) %>% 
  group_rows("Diagnosis",13,17,bold=FALSE) %>% 
  collapse_rows(columns = 5,valign = "top", latex_hline = "none") %>% 
  footnote(general = "Continuous variables expressed as median (Q1,Q3), Wilcoxon rank sum test; Categorical variables expressed as N(%), Chi square/*Fisher's",threeparttable=T)



kable(df.tab[18:nrow(df.tab),], "latex", booktabs = T, align = "l", col.names = c("All patients (n=64)", "PRA- (n=35)", "PRA+ (n=29)", "p value"), caption = "Table 1b: Intra and Postoperative Variables") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Intraoperative",1,6,bold=TRUE) %>% 
  group_rows("ABO incompatible",3,4,bold=FALSE) %>% 
  group_rows("Plasmapheresis",5,6,bold=FALSE) %>% 
  group_rows("Postoperative",7,15,bold=TRUE) %>% 
  group_rows("Rejection",10,11,bold=FALSE) %>% 
  group_rows("Donor specific antibody",12,13,bold=FALSE) %>% 
  group_rows("ECMO",14,15,bold=FALSE) %>% 
  collapse_rows(columns = 5,valign = "top", latex_hline = "none") %>% 
  footnote(general = "Continuous variables expressed as median (Q1,Q3), Wilcoxon rank sum test; Categorical variables expressed as N(%), Chi square/*Fisher's",threeparttable=T)

```


\clearpage

```{r bloodprodclean}
##create binary variables for did/did not have prior exposures
df.blood <- df.all[!is.na(df.all$prior_donor_exposures),]
df.blood$prior_bin <- factor(ifelse(df.blood$prior_donor_exposures>0,1,0),levels = c(0,1))
df.blood$rbc_bin <- factor(ifelse(df.blood$prior_rbc>0,1,0),levels = c(0,1))
df.blood$ffp_bin <- factor(ifelse(df.blood$prior_ffp>0,1,0),levels = c(0,1))
df.blood$plt_bin <- factor(ifelse(df.blood$prior_plt>0,1,0),levels = c(0,1))
df.blood$cryo_bin <- factor(ifelse(df.blood$prior_cryo>0,1,0),levels = c(0,1))
df.blood$cell <- df.blood$prior_rbc + df.blood$prior_plt
df.blood$acell <- df.blood$prior_ffp + df.blood$prior_cryo
df.blood$cell_bin <- factor(ifelse(df.blood$cell>0,1,0),levels = c(0,1))
df.blood$acell_bin <- factor(ifelse(df.blood$acell>0,1,0),levels = c(0,1))
df.blood$pra_bin <- factor(ifelse(df.blood$pra_status==1,0,1),levels=c(0,1))

```


```{r studyobjectives}

mystats2 <- function(X,Y){
  if(class(X)=="numeric"&&class(Y)=="factor"){
    tab <- matrix(NA,1,3)
    all_1 <- format(round(median(X,na.rm = TRUE),1),nsmall = 1)
    all_2 <- paste0("(",format(round(quantile(X,na.rm = TRUE)[2],1),nsmall=1),", ",
                    format(round(quantile(X,na.rm = TRUE)[4],1),nsmall=1),")")
    
    strat1_1 <- format(round(median(X[Y==levels(Y)[1]],na.rm = TRUE),1),nsmall=1)
    strat2_1 <- format(round(median(X[Y==levels(Y)[2]],na.rm = TRUE),1),nsmall=1)
    
    strat1_2 <- paste0("(",format(round(quantile(X[Y==levels(Y)[1]],na.rm = TRUE)[2],1),nsmall=1),", ",
                       format(round(quantile(X[Y==levels(Y)[1]],na.rm = TRUE)[4],1),nsmall=1),")")
    strat2_2 <- paste0("(",format(round(quantile(X[Y==levels(Y)[2]],na.rm = TRUE)[2],1),nsmall = 1),", ",
                       format(round(quantile(X[Y==levels(Y)[2]],na.rm = TRUE)[4],1),nsmall = 1),")")
    
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)
  }
  
  if(class(X)=="factor"&&class(Y)=="factor"){
        nlev <- nlevels(X)
    all_1 <- c()
    all_2 <- c()
    strat1_1 <- c()
    strat2_1 <- c()
    strat1_2 <- c()
    strat2_2 <- c()
    tab <- matrix(NA,nlev,3)
    for(i in 1:nlev){
    tab1 <- table(X)
    all_1[i] <- paste(tab1[i])
    all_2[i] <- paste0("(",format(round(prop.table(tab1)[i]*100,1),nsmall = 1),")")
    
    tab2 <- table(X,Y)
    strat1_1[i] <- tab2[i,1]
    strat1_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,1]*100,1),nsmall = 1),")")
    
    strat2_1[i] <- tab2[i,2]
    strat2_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,2]*100,1),nsmall = 1),")")
      
    }
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)

  }
  if(class(X)=="numeric"&&class(Y)=="numeric"){
    
    tab <- cor.test(df.blood$prior_donor_exposures,df.blood$cpra)$estimate

  }

  return(tab)
}

varlist <- c("prior_bin","rbc_bin","plt_bin","ffp_bin","cryo_bin","prior_donor_exposures","prior_rbc","prior_plt","prior_ffp","prior_cryo")

myfulltab <- lapply(df.blood[varlist], FUN = mystats2, Y = df.blood$pra_bin)

mat.tab <- as.matrix(unlist(myfulltab[[1]]))
for (i in 2:length(myfulltab)){
  mat.tab <- rbind(mat.tab,as.matrix(unlist(myfulltab[[i]])))
}

mypvals <- lapply(df.blood[varlist], function(x)
                  rep(mypval(x,df.blood$pra_bin)$pval,varlevels(x)))
mypvalmat <- as.matrix(unlist(mypvals))
mat.tab.p <- cbind(mat.tab,mypvalmat)
df.tab <- data.frame(mat.tab.p)

otherstats <- function(X,Y){
  if(class(X)=="factor"&&class(Y)=="numeric"){
  stattab <- t(mystats2(Y,X))
  pval <- rep(format(round(wilcox.test(Y[X==levels(X)[1]],Y[X==levels(X)[2]])$p.value,2),nsmall=2),nlevels(X))
  tab <- cbind(stattab,pval)
  }
  if(class(X)=="numeric"&&class(Y)=="numeric"){
      stattab <- format(round(cor.test(X,Y,method = "spearman")$estimate,2),nsmall=2)
      pval <- format(round(cor.test(X,Y,method = "spearman")$p.value,2),nsmall = 2)
      tab <- cbind(stattab,pval)
  }
  
  return(tab)
}


varlist2 <- c("prior_bin","rbc_bin","plt_bin","ffp_bin","cryo_bin","prior_donor_exposures","prior_rbc","prior_plt","prior_ffp","prior_cryo")

myfulltab2 <- lapply(df.blood[varlist], FUN = otherstats, Y = df.blood$cpra)

mat.tab2 <- as.matrix(unlist(myfulltab2[[1]]))
for (i in 2:length(myfulltab2)){
  mat.tab2 <- rbind(mat.tab2,as.matrix(unlist(myfulltab2[[i]])))
}


df.tab2 <- cbind(insertRows(df.tab,c(1,4,7,10,13),as.data.frame(matrix("",5,ncol(df.tab)))),data.frame(mat.tab2))

rownames(df.tab2) <- c("Total prior donor exposures","No","Yes","Red Blood Cells","No ","Yes ","Platelets","No  ","Yes  ","Fresh frozen plasma","No   ","Yes   ","Cryoprecipitate","No    ","Yes    ","Total prior donor exposures ","Red blood cells ","Platelets ","Fresh frozen plasma ","Cryoprecipitate ")

kable(df.tab2[,1:4], "latex", booktabs = T, align = "l", col.names = c("All patients (n=64)", "PRA- (n=35)", "PRA+ (n=29)", "p value"), caption = "Table 2a: Blood products and PRA (binary)") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Blood products - binary",1,15,bold=TRUE) %>% 
  group_rows("Blood products - count", 16, 20, bold = TRUE) %>% 
  collapse_rows(columns = 5,valign = "top", latex_hline = "none") %>% 
  row_spec(15, hline_after = TRUE) %>%
  add_indent(c(2,3,5,6,8,9,11,12,14,15)) %>% 
  footnote(general = "Continuous variables: median (Q1,Q3), Wilcoxon rank sum test; Categorical variables: N(%), Chi square/Fisher's",threeparttable=T)


kable(df.tab2[,5:6], "latex", booktabs = T, align = "l", col.names = c("cPRA", "p value"), caption = "Table 2b: Blood products and cPRA (continuous)") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Blood products - binary",1,15,bold=TRUE) %>% 
  group_rows("Blood products - count", 16, 20, bold = TRUE) %>% 
  collapse_rows(columns = 3,valign = "top", latex_hline = "none") %>% 
  add_indent(c(2,3,5,6,8,9,11,12,14,15)) %>% 
  row_spec(15, hline_after = TRUE) %>%
  footnote(general = "Binary blood products expressed as median (Q1,Q3), Wilcoxon rank sum test; Count blood products expressed as Spearman rank correlation $r_s$",threeparttable=T,escape=F)

```


\clearpage


```{r bloodprodtabs}
make2x2table <- function(X,Y,mycols,myrows,caption){
  tab <- my2x2(X,Y)
  pval <- mypval(X,Y,FALSE)
  foot <- paste0("p = ",pval$pval," (",pval$test,")")
  rownames(tab) <- myrows
  rowspec <- nlevels(X)
  mytable <- kable(tab, col.names = mycols, row.names = TRUE, "latex", booktabs = T,
                   caption=caption) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE) %>%
  footnote(general = foot) %>%
  row_spec(rowspec, hline_after = TRUE) %>%
  column_spec(c(2,4), border_left  = TRUE)

  return(mytable)
}
colnames.all <- c("PRA-", "PRA+", "Total")

```

```{r alleles}
df.alleles <- df.blood[df.blood$pra_status==2,]

df.alleles$allele_c <- ifelse(df.alleles$class_i==1 & df.alleles$class_ii==1,3,
                              ifelse(df.alleles$class_i==1 & df.alleles$class_ii==0,1,2))

df.alleles$allele_c <- factor(df.alleles$allele_c, levels=c(1,2,3),
                              labels = c("1","2","both"))

```

```{r tab17}
make2x2table(df.alleles$allele_c,df.alleles$donor_c,
             mycols = c("No donor \nspecific antibody","Donor specific \nantibody","Total"),
             myrows=c("Class I","Class II","Class I and II","Total"),
             "Table 3: Allele class and donor specific antibody")
```


\section{Questions/Decisions to make as we move into modeling}

1. Should blood products be modeled as **binary** or **counts**?
2. Should individual blood products be part of the model or only the total?
3. Is it sufficient to only control for demographic/preoperative variables?
    - One reviewer states: "the association of PRA and outcomes should be adjusted to covariates known to affect all; time to extubation, ICU length of stay and total length of stay" but these outcomes occur **after** PRA and blood products are measured, so to me it does not make sense to control for them in the model. We should discuss as a group.
4. We will need to condense **race** and **diagnosis** because some categories have very low counts. What is the most logical way to group each of these variables? 

\newpage

\section{Univariate Results}

```{r univar}
unimod <- function(var){
    mod <- glm(pra_bin ~ var,data=df.blood,family="binomial")
    coefs <- round(summary(mod)$coefficients[,1],2)
    coefs.p <- matrix(paste0(coefs," (",round(summary(mod)$coefficients[,2],2),")"),nrow=2,ncol=1)
    oddsrat <- matrix(paste0(round(exp(coefs),2)," (",round(exp(confint(mod))[,1],2),", ",round(exp(confint(mod))[,2],2),")"),nrow=2,ncol=1)
    pval <- matrix(rep(round(summary(mod)$coefficients[2,4],3),2),nrow=2,ncol=1)
    
    tab <- cbind(coefs.p,oddsrat,pval)

return(tab)
}


varlist <- c("prior_bin","rbc_bin","plt_bin","ffp_bin","cryo_bin","prior_donor_exposures","prior_rbc","prior_plt","prior_ffp","prior_cryo")

modlist <- lapply(df.blood[varlist], FUN=unimod)

mat.tab <- as.matrix(unlist(modlist[[1]]))
for (i in 2:length(modlist)){
  mat.tab <- rbind(mat.tab,as.matrix(unlist(modlist[[i]])))
}

rownames(mat.tab) <- c("Intercept","Total","Intercept ","Red blood cells","Intercept  ","Platelets","Intercept   ","Fresh frozen plasma","Intercept    ","Cryoprecipitate","Intercept     ","Total ","Intercept      ","Red blood cells ","Intercept       ","Platelets ","Intercept        ","Fresh frozen plasma ","Intercept         ","Cryoprecipitate ")

kable(mat.tab, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","OR (95% CI)" ,"p value"), caption = "Table 4: Univariate results for blood product exposures") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Blood products - binary",1,10,bold=TRUE) %>% 
  group_rows("Blood products - count", 11, 20, bold = TRUE) %>% 
  row_spec(c(2,4,6,8,10,12,14,16,18), hline_after = TRUE) %>%
  collapse_rows(columns = 4,valign = "bottom", latex_hline = "none") %>% 
  footnote(general = "Absence of blood product is the refernce group for all binary variables")



cormat <- matrix(cor(df.blood[,c("prior_rbc","prior_plt","prior_ffp","prior_cryo")]),4,4)
rownames(cormat) <- c("RBCs","Plt","FFP","Cryo")
kable(cormat, "latex", booktabs = T, align = "l", col.names = c("RBCs","Plt","FFP","Cryo"), caption = "Table 5: Correlation matrix for blood products") %>%
  kable_styling(latex_options = "hold_position")


demvarlist <- c("age","weight","no_of_prior_surgeries","gender_c","prior_mech_nomiss")

modlist2 <- lapply(df.blood[demvarlist], FUN=unimod)

mat.tab2 <- as.matrix(unlist(modlist2[[1]]))
for (i in 2:length(modlist2)){
  mat.tab2 <- rbind(mat.tab2,as.matrix(unlist(modlist2[[i]])))
}

mat.tab2r <- insertRows(as.data.frame(mat.tab2),8,c("--","--",mat.tab2[7,3]))
mat.tab2r <- insertRows(as.data.frame(mat.tab2r),11,c("--","--",mat.tab2[10,3]))

rownames(mat.tab2r) <- c("Intercept","Age","Intercept ","Weight","Intercept  ","Number of previous surgeries","Intercept   ","Male (ref)","Female","Intercept    ","No (ref)","Yes")

kable(mat.tab2r, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","OR (95% CI)" ,"p value"), caption = "Table 7: Univariate results for other demographic variables") %>%
  kable_styling(latex_options = "hold_position") %>%
  group_rows("Sex",7,9,bold=FALSE) %>%
  group_rows("Prior mechanical support", 10, 12, bold = FALSE) %>%
  row_spec(c(2,4,6,9), hline_after = TRUE) %>%
  collapse_rows(columns = 4,valign = "bottom", latex_hline = "none")


```


```{r model, eval=FALSE}
df.blood$race2 <- factor(ifelse(df.blood$race_c=="Hispanic"|df.blood$race_c=="Asian"|df.blood$race_c=="No response","H/A/NR",df.blood$race_c))

mod <- glm(pra_bin ~ prior_bin + age + weight + prior_mech_nomiss +  no_of_prior_surgeries + gender_c + race2 + diag_c, data=df.blood, family="binomial",na.action = na.pass)

#step.aic <- step(mod)

finalmod <- glm(pra_bin ~ prior_bin + prior_mech_nomiss + race2 + diag_c,data=df.blood,family="binomial")


coefs <- summary(finalmod)$coefficients[,1]
coefs.se <- summary(finalmod)$coefficients[,2]
oddsrat <- exp(coefs)
or.se <- exp(confint(finalmod))
pvals <- summary(finalmod)$coefficients[,4]

mymat <- round(cbind(coefs,coefs.se,oddsrat,or.se,pvals),3)

tabmat <- matrix(0,nrow(mymat),3)
tabmat[,1] <- paste0(mymat[,1]," (",mymat[,2],")")
tabmat[,2] <- paste0(mymat[,3]," (",mymat[,4],",",mymat[,5],")")
tabmat[,3] <- mymat[,6]

rownames(tabmat) <- c("Intercept","Total blood products","Mech support","Race (AA)","Race (H/A/NR)","Diagnosis (Dilated CM)","Diagnosis (Failed SV)","Diagnosis (Restrictive CM)","Diagnosis (Other)")
kable(tabmat, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","Odds Ratio (95% CI)","p value"), caption = "Final Model") %>%
  kable_styling(latex_options = "hold_position")

# library(pROC)
# test_prob = predict(finalmod, type = "response")
# test_roc = roc(df.blood$pra_bin ~ test_prob, plot = TRUE, print.auc = TRUE)

```

```{r modelparts, eval=FALSE}
mod2 <- glm(pra_bin ~ rbc_bin + plt_bin + ffp_bin + cryo_bin + prior_mech_nomiss + race2 + diag_c, data=df.blood, family="binomial",na.action = na.pass)

coefs <- summary(mod2)$coefficients[,1]
coefs.se <- summary(mod2)$coefficients[,2]
oddsrat <- exp(coefs)
or.se <- exp(confint(mod2))
pvals <- summary(mod2)$coefficients[,4]

mymat <- round(cbind(coefs,coefs.se,oddsrat,or.se,pvals),3)

tabmat <- matrix(0,nrow(mymat),3)
tabmat[,1] <- paste0(mymat[,1]," (",mymat[,2],")")
tabmat[,2] <- paste0(mymat[,3]," (",mymat[,4],",",mymat[,5],")")
tabmat[,3] <- mymat[,6]

rownames(tabmat) <- c("Intercept","Red blood cells","Platelets","Fresh frozen plasma","Cryoprecipitate","Mech support","Race (AA)","Race (H/A/NR)","Diagnosis (Dilated CM)","Diagnosis (Failed SV)","Diagnosis (Restrictive CM)","Diagnosis (Other)")
kable(tabmat, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","Odds Ratio (95% CI)","p value"), caption = "Model with Individual Blood Components") %>%
  kable_styling(latex_options = "hold_position")

test_prob2 = predict(mod2, type = "response")
test_roc2 = roc(df.blood$pra_bin ~ test_prob2, plot = TRUE, print.auc = TRUE)



```



