---
title: "Panel Reactive Antibodies (PRAs) and Their Association with Blood Product Exposures in Pediatric Patients Undergoing Cardiac Transplantation"
author:
  - 'Dr. Nina Guzzetta, Dr. Susana Cruz Beltran'
  - 'Statisticians: Dr. Renee Moore, Andrea Lane'
date: "3/10/2021"
geometry: margin=1cm
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
library(readxl)
library(epitrix)
library(dplyr)
library(rockchalk)
library(kableExtra)
library(summarytools)
library(ggplot2)
library(ggmosaic)
library(tidyverse)
library(berryFunctions)
library(caret)
library(glmnet)

options(knitr.table.format = "latex")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00")

```

```{r upload}
df.sheet1 <- read_excel("Transplant data.final.xlsx", sheet = 1,
                        col_types = c("text","text","date","text","numeric",
                                      "text","text","text", "date","date",
                                      "text","text","text","numeric","text",
                                      "text",rep("numeric",5),"text",rep("numeric",16)))
df.sheet2 <- read_excel("Transplant data.final.xlsx", sheet = 2,
                        col_types = c("text","text","date","text","numeric",
                                      "text","text","text", "date","date",
                                      "text","text","text","numeric",rep("numeric",6),"text",
                                      "text",rep("numeric",5),"text",rep("numeric",16)))

df.neg <- df.sheet1[-1,]
df.pos <- df.sheet2[-1,]

names(df.neg) <- clean_labels(names(df.neg))
names(df.pos) <- clean_labels(names(df.pos))

names(df.neg)[1] <- "Subj_num"
names(df.pos)[1] <- "Subj_num"

names(df.neg)[25] <- "prior_rbc"
names(df.neg)[26] <- "prior_ffp"
names(df.neg)[27] <- "prior_plt"
names(df.neg)[28] <- "prior_cryo"

names(df.neg)[30] <- "dos_rbc"
names(df.neg)[31] <- "dos_ffp"
names(df.neg)[32] <- "dos_plt"
names(df.neg)[33] <- "dos_cryo"

names(df.neg)[35] <- "post_rbc"
names(df.neg)[36] <- "post_ffp"
names(df.neg)[37] <- "post_plt"
names(df.neg)[38] <- "post_cryo"


names(df.pos)[31] <- "prior_rbc"
names(df.pos)[32] <- "prior_ffp"
names(df.pos)[33] <- "prior_plt"
names(df.pos)[34] <- "prior_cryo"

names(df.pos)[36] <- "dos_rbc"
names(df.pos)[37] <- "dos_ffp"
names(df.pos)[38] <- "dos_plt"
names(df.pos)[39] <- "dos_cryo"

names(df.pos)[41] <- "post_rbc"
names(df.pos)[42] <- "post_ffp"
names(df.pos)[43] <- "post_plt"
names(df.pos)[44] <- "post_cryo"

df.all <- bind_rows("1" = df.neg, "2" = df.pos, .id="pra_status")

```

```{r clean}
df.all$age_unit <- substr(df.all$age_years,nchar(df.all$age_years),nchar(df.all$age_years))
df.all$age_num <- substr(df.all$age_years,1,nchar(df.all$age_years)-1)

df.all$age <- ifelse(df.all$age_unit == "m", as.numeric(df.all$age_num)/12, ifelse(df.all$age_unit == "w", as.numeric(df.all$age_num)/52, as.numeric(df.all$age_num)))

df.all$abo_incompatible_nomiss <- ifelse(is.na(df.all$abo_incompatible),0,1)
df.all$prior_mech_nomiss <- ifelse(is.na(df.all$prior_mech_support),0,1)

df.all$plasma_c <- ifelse(df.all$plasmapheresis %in% c("Yes","yes"),1,0)
df.all$rej_c <- ifelse(df.all$rejection=="N",0,1)

df.all$plasma_c <- factor(df.all$plasma_c, levels = c(0,1),
                          labels = c("No","Yes"))
df.all$rej_c <- factor(df.all$rej_c, levels = c(0,1), labels = c("No","Yes"))
df.all$ecmo_c <- factor(df.all$ecmo_support_postop, levels = c("no","yes"),
                        labels = c("no","yes"))
df.all$diag_c <- df.all$diagnosis
df.all$diag_c[df.all$diag_c=="LV failure"] <- "Dilated CM"
df.all$diag_c[df.all$diag_c=="RV Failure" | df.all$diag_c=="Shone's complex- Coarct - Multiple VSD"] <- "Other"

df.all$diag_c <- factor(df.all$diag_c, 
      levels = c("ARVD","Dilated CM","Failed SV","Restrictive CM","Other"),
      labels = c("ARVD","Dilated CM","Failed SV","Restrictive CM","Other"))
df.all$donor_c <- factor(df.all$donor_specific_ab, levels = c("N","Y"),
                         labels = c("No","Yes"))
df.all$gender_c <- factor(df.all$gender, levels = c("F","M"), labels = c("Female","Male"))
df.all$race_c <- factor(df.all$race, 
        levels = c("White","African American","Hispanic","Asian","Declined"),             labels = c("White","African American","Hispanic","Asian","No response"))
df.all$abo_incompatible_nomiss <- factor(df.all$abo_incompatible_nomiss, 
                                         levels = c(0,1), labels = c("No","Yes"))
df.all$prior_mech_nomiss <- factor(df.all$prior_mech_nomiss, levels = c(0,1),
                                         labels = c("No","Yes"))

df.all <- df.all[!is.na(df.all$prior_donor_exposures),]
```

```{r functions}
my2x2 <- function(X,Y,col=TRUE){
  perc <- ifelse(col,2,1)
  mytab <- table(X,Y)
  proptab <- format(round(prop.table(mytab,perc)*100,1),nsmall = 1,trim=TRUE)
  mytab2 <- matrix(paste0(mytab," (",proptab,")"),nlevels(X),2)
  mytab2 <- rbind(mytab2, colSums(mytab))
  mytab <- cbind(mytab2, c(rowSums(mytab),sum(rowSums(mytab))))
  return(mytab)
}

mypval <- function(X,Y,symbol=TRUE){
  if (class(X)=="numeric"){
    test <- "Wilcoxon rank-sum"
    pval <- wilcox.test(X[df.all$pra_status==1],X[df.all$pra_status==2])$p.value
    pval <- ifelse(pval == 1,0.99,pval)
    pval <- format(round(pval,2), nsmall=2)
    pval <- ifelse(pval==0,"<0.0001",pval)
  } else if (class(X)=="factor"){
  
  tab <- table(X,Y)
  exp.tab <- chisq.test(tab)$expected
  test <- ifelse(all(exp.tab > 5),"Chi Square","Fisher's Exact")
  pval <- ifelse(all(exp.tab > 5),chisq.test(tab,correct = FALSE)$p.value,fisher.test(tab)$p.value)
  pval <- ifelse(pval == 1,0.99,pval)
  pval <- format(round(pval,2), nsmall=2)
  pval <- ifelse(pval=="0.00","<0.01",pval)
  if(symbol){pval <- ifelse(test=="Chi Square",
                 paste0(pval,""),
                 paste0(pval,"*"))}
   }
  return(list(pval = pval, test = test))
}

varlevels <- function(X){
  if(class(X)=="numeric"){
    nlev <- 1
  } else if (class(X)=="factor"){
    nlev <- nlevels(X)
  }
  return(nlev)
}

mystats <- function(X,Y){
  if (class(X)=="numeric"){
    tab <- matrix(NA,1,3)
    all_1 <- format(round(median(X,na.rm = TRUE),1),nsmall = 1)
    all_2 <- paste0("(",format(round(quantile(X,na.rm = TRUE)[2],1),nsmall=1),",",
                    format(round(quantile(X,na.rm = TRUE)[4],1),nsmall=1),")")
    
    strat1_1 <- format(round(median(X[Y==1],na.rm = TRUE),1),nsmall=1)
    strat2_1 <- format(round(median(X[Y==2],na.rm = TRUE),1),nsmall=1)
    
    strat1_2 <- paste0("(",format(round(quantile(X[Y==1],na.rm = TRUE)[2],1),nsmall=1),",",
                       format(round(quantile(X[Y==1],na.rm = TRUE)[4],1),nsmall=1),")")
    strat2_2 <- paste0("(",format(round(quantile(X[Y==2],na.rm = TRUE)[2],1),nsmall = 1),",",
                       format(round(quantile(X[Y==2],na.rm = TRUE)[4],1),nsmall = 1),")")
    
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)
    
    tab <- cbind(cell1,cell2,cell3)
  } else if (class(X)=="factor"){
    nlev <- nlevels(X)
    all_1 <- c()
    all_2 <- c()
    strat1_1 <- c()
    strat2_1 <- c()
    strat1_2 <- c()
    strat2_2 <- c()
    tab <- matrix(NA,nlev,3)
    for(i in 1:nlev){
    tab1 <- table(X)
    all_1[i] <- paste(tab1[i])
    all_2[i] <- paste0("(",format(round(prop.table(tab1)[i]*100,1),nsmall = 1),")")
    
    tab2 <- table(X,Y)
    strat1_1[i] <- tab2[i,1]
    strat1_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,1]*100,1),nsmall = 1),")")
    
    strat2_1[i] <- tab2[i,2]
    strat2_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,2]*100,1),nsmall = 1),")")
      
    }
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)

  }
  return(tab)
}

```

\clearpage
\section{Tables}
```{r table1}
varlist <- c("age","weight","no_of_prior_surgeries","gender_c","race_c","prior_mech_nomiss","diag_c","donor_time","cpb_time","abo_incompatible_nomiss","plasma_c","icu_length_of_stay_day", "vis", "time_to_extubation_hr","rej_c", "donor_c","ecmo_c")
# var_levels <- lapply(df.all[varlist], FUN = varlevels)
# df_levels <- data.frame(matrix(unlist(var_levels), nrow=length(varlist),byrow = T))

myfulltab <- lapply(df.all[varlist], FUN = mystats, Y = df.all$pra_status)

mat.tab <- as.matrix(unlist(myfulltab[[1]]))
for (i in 2:length(myfulltab)){
  mat.tab <- rbind(mat.tab,as.matrix(unlist(myfulltab[[i]])))
}

mypvals <- lapply(df.all[varlist], function(x)
                  rep(mypval(x,df.all$pra_status)$pval,varlevels(x)))
# mytest <- lapply(df.all[varlist], function(x)
#                   rep(mypval(x,df.all$pra_status)$test,varlevels(x)))
# mytestmat <- as.matrix(unlist(mytest))
mypvalmat <- as.matrix(unlist(mypvals))
mat.tab.p <- cbind(mat.tab,mypvalmat)
df.tab <- data.frame(mat.tab.p)

names(df.tab)[2] <- paste0(names(df.tab)[2],
  footnote_marker_symbol(1, "latex"))
rownames(df.tab) <- c("Age (Years)", "Weight at transplantation (kg)", 
                      "Number of previous surgeries","Female", "Male", "White",
                      "African American", "Hispanic","Asian", "No response",  
                       "No", "Yes","ARVD","Dilated CM","Failed SV",
                      "Restrictive CM","Other","Donor cross clamp time (min)",
                      "CPB time (min)","No ","Yes ","No  ","Yes  ",
                      "Length of ICU stay (days)","Vasoactive inotrope score",
                      "Time to extubation (hrs)","No   ","Yes   ","No    ",
                      "Yes    ","No     ","Yes     ")

kable(df.tab, "latex", booktabs = T, align = "l", col.names = c("All patients (n=64)", "PRA- (n=35)", "PRA+ (n=29)", "p value"), caption = "Pre, Intra, and Postoperative Variables") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Demographic and preoperative",1,17,bold=TRUE) %>% 
  group_rows("Gender", 4, 5, bold = FALSE) %>% 
  group_rows("Race", 6, 10, bold = FALSE) %>% 
  group_rows("Mechanical support prior to transplantation", 11, 12, bold = FALSE) %>% 
  group_rows("Intraoperative",18,23,bold=TRUE) %>% 
  group_rows("Diagnosis",13,17,bold=FALSE) %>% 
  group_rows("ABO incompatible",20,21,bold=FALSE) %>% 
  group_rows("Plasmapheresis",22,23,bold=FALSE) %>% 
  group_rows("Postoperative",24,32,bold=TRUE) %>% 
  group_rows("Rejection",27,28,bold=FALSE) %>% 
  group_rows("Donor specific antibody",29,30,bold=FALSE) %>% 
  group_rows("ECMO",31,32,bold=FALSE) %>% 
  collapse_rows(columns = 5,valign = "top", latex_hline = "none") %>% 
  footnote(general = "Continuous variables: median (Q1,Q3), Wilcoxon rank sum test; Categorical variables: N(%), Chi square/Fisher's")

```


\clearpage

```{r bloodprodclean}
##create binary variables for did/did not have prior exposures
df.blood <- df.all[!is.na(df.all$prior_donor_exposures),]
df.blood$prior_bin <- factor(ifelse(df.blood$prior_donor_exposures>0,1,0),levels = c(0,1))
df.blood$rbc_bin <- factor(ifelse(df.blood$prior_rbc>0,1,0),levels = c(0,1))
df.blood$ffp_bin <- factor(ifelse(df.blood$prior_ffp>0,1,0),levels = c(0,1))
df.blood$plt_bin <- factor(ifelse(df.blood$prior_plt>0,1,0),levels = c(0,1))
df.blood$cryo_bin <- factor(ifelse(df.blood$prior_cryo>0,1,0),levels = c(0,1))
df.blood$cell <- df.blood$prior_rbc + df.blood$prior_plt
df.blood$acell <- df.blood$prior_ffp + df.blood$prior_cryo
df.blood$cell_bin <- factor(ifelse(df.blood$cell>0,1,0),levels = c(0,1))
df.blood$acell_bin <- factor(ifelse(df.blood$acell>0,1,0),levels = c(0,1))
df.blood$pra_bin <- factor(ifelse(df.blood$pra_status==1,0,1),levels=c(0,1))

```


```{r studyobjectives}

mystats2 <- function(X,Y){
  if(class(X)=="numeric"&&class(Y)=="factor"){
    tab <- matrix(NA,1,3)
    all_1 <- format(round(median(X,na.rm = TRUE),1),nsmall = 1)
    all_2 <- paste0("(",format(round(quantile(X,na.rm = TRUE)[2],1),nsmall=1),",",
                    format(round(quantile(X,na.rm = TRUE)[4],1),nsmall=1),")")
    
    strat1_1 <- format(round(median(X[Y==levels(Y)[1]],na.rm = TRUE),1),nsmall=1)
    strat2_1 <- format(round(median(X[Y==levels(Y)[2]],na.rm = TRUE),1),nsmall=1)
    
    strat1_2 <- paste0("(",format(round(quantile(X[Y==levels(Y)[1]],na.rm = TRUE)[2],1),nsmall=1),",",
                       format(round(quantile(X[Y==levels(Y)[1]],na.rm = TRUE)[4],1),nsmall=1),")")
    strat2_2 <- paste0("(",format(round(quantile(X[Y==levels(Y)[2]],na.rm = TRUE)[2],1),nsmall = 1),",",
                       format(round(quantile(X[Y==levels(Y)[2]],na.rm = TRUE)[4],1),nsmall = 1),")")
    
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)
  }
  
  if(class(X)=="factor"&&class(Y)=="factor"){
        nlev <- nlevels(X)
    all_1 <- c()
    all_2 <- c()
    strat1_1 <- c()
    strat2_1 <- c()
    strat1_2 <- c()
    strat2_2 <- c()
    tab <- matrix(NA,nlev,3)
    for(i in 1:nlev){
    tab1 <- table(X)
    all_1[i] <- paste(tab1[i])
    all_2[i] <- paste0("(",format(round(prop.table(tab1)[i]*100,1),nsmall = 1),")")
    
    tab2 <- table(X,Y)
    strat1_1[i] <- tab2[i,1]
    strat1_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,1]*100,1),nsmall = 1),")")
    
    strat2_1[i] <- tab2[i,2]
    strat2_2[i] <- paste0("(",format(round(prop.table(tab2,2)[i,2]*100,1),nsmall = 1),")")
      
    }
    cell1 <- paste(all_1,all_2)
    cell2 <- paste(strat1_1,strat1_2)
    cell3 <- paste(strat2_1,strat2_2)

    tab <- cbind(cell1,cell2,cell3)

  }
  if(class(X)=="numeric"&&class(Y)=="numeric"){
    
    tab <- cor.test(df.blood$prior_donor_exposures,df.blood$cpra)$estimate

  }

  return(tab)
}

varlist <- c("prior_bin","rbc_bin","plt_bin","ffp_bin","cryo_bin","prior_donor_exposures","prior_rbc","prior_plt","prior_ffp","prior_cryo")

myfulltab <- lapply(df.blood[varlist], FUN = mystats2, Y = df.blood$pra_bin)

mat.tab <- as.matrix(unlist(myfulltab[[1]]))
for (i in 2:length(myfulltab)){
  mat.tab <- rbind(mat.tab,as.matrix(unlist(myfulltab[[i]])))
}

mypvals <- lapply(df.blood[varlist], function(x)
                  rep(mypval(x,df.blood$pra_bin)$pval,varlevels(x)))
mypvalmat <- as.matrix(unlist(mypvals))
mat.tab.p <- cbind(mat.tab,mypvalmat)
df.tab <- data.frame(mat.tab.p)

otherstats <- function(X,Y){
  if(class(X)=="factor"&&class(Y)=="numeric"){
  stattab <- t(mystats2(Y,X))
  pval <- rep(format(round(wilcox.test(Y[X==levels(X)[1]],Y[X==levels(X)[2]])$p.value,2),nsmall=2),nlevels(X))
  tab <- cbind(stattab,pval)
  }
  if(class(X)=="numeric"&&class(Y)=="numeric"){
      stattab <- format(round(cor.test(X,Y,method = "spearman")$estimate,2),nsmall=2)
      pval <- format(round(cor.test(X,Y,method = "spearman")$p.value,2),nsmall = 2)
      tab <- cbind(stattab,pval)
  }
  
  return(tab)
}


varlist2 <- c("prior_bin","rbc_bin","plt_bin","ffp_bin","cryo_bin","prior_donor_exposures","prior_rbc","prior_plt","prior_ffp","prior_cryo")

myfulltab2 <- lapply(df.blood[varlist], FUN = otherstats, Y = df.blood$cpra)

mat.tab2 <- as.matrix(unlist(myfulltab2[[1]]))
for (i in 2:length(myfulltab2)){
  mat.tab2 <- rbind(mat.tab2,as.matrix(unlist(myfulltab2[[i]])))
}


df.tab2 <- cbind(insertRows(df.tab,c(1,4,7,10,13),as.data.frame(matrix("",5,ncol(df.tab)))),data.frame(mat.tab2))

rownames(df.tab2) <- c("Total prior donor exposures","No","Yes","Red Blood Cells","No ","Yes ","Platelets","No  ","Yes  ","Fresh frozen plasma","No   ","Yes   ","Cryoprecipitate","No    ","Yes    ","Total prior donor exposures ","Red blood cells ","Platelets ","Fresh frozen plasma ","Cryoprecipitate ")

kable(df.tab2, "latex", booktabs = T, align = "l", col.names = c("All patients (n=64)", "PRA- (n=35)", "PRA+ (n=29)", "p value", "cPRA","p value"), caption = "Blood products and PRA") %>%
  kable_styling(latex_options = "hold_position") %>% 
  group_rows("Blood products - binary",1,15,bold=TRUE) %>% 
  group_rows("Blood products - count", 16, 20, bold = TRUE) %>% 
  collapse_rows(columns = 5,valign = "top", latex_hline = "none") %>% 
  collapse_rows(columns = 7,valign = "top", latex_hline = "none") %>% 
  add_header_above(c(" "=1,"PRA binary"=4,"cPRA continuous"=2)) %>% 
  row_spec(15, hline_after = TRUE) %>%
  column_spec(6, border_left  = TRUE) %>%
  footnote(general = "Continuous variables: median (Q1,Q3), Wilcoxon rank sum test; Categorical variables: N(%), Chi square/Fisher's")

```


\clearpage


```{r bloodprodtabs}
make2x2table <- function(X,Y,mycols,myrows,caption){
  tab <- my2x2(X,Y)
  pval <- mypval(X,Y,FALSE)
  foot <- paste0("p = ",pval$pval," (",pval$test,")")
  rownames(tab) <- myrows
  rowspec <- nlevels(X)
  mytable <- kable(tab, col.names = mycols, row.names = TRUE, "latex", booktabs = T,
                   caption=caption) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE) %>%
  footnote(general = foot) %>%
  row_spec(rowspec, hline_after = TRUE) %>%
  column_spec(c(2,4), border_left  = TRUE)

  return(mytable)
}
colnames.all <- c("PRA-", "PRA+", "Total")

```

```{r alleles}
df.alleles <- df.blood[df.blood$pra_status==2,]

df.alleles$allele_c <- ifelse(df.alleles$class_i==1 & df.alleles$class_ii==1,3,
                              ifelse(df.alleles$class_i==1 & df.alleles$class_ii==0,1,2))

df.alleles$allele_c <- factor(df.alleles$allele_c, levels=c(1,2,3),
                              labels = c("1","2","both"))

```

```{r tab17}
make2x2table(df.alleles$allele_c,df.alleles$donor_c,
             mycols = c("No donor \nspecific antibody","Donor specific \nantibody","Total"),
             myrows=c("Class I","Class II","Class I and II","Total"),
             "Allele class and donor specific antibody")
```
Table 3 includes only PRA+ patients and explores the relationship between allele class and development of a donor specific antibody. 100% of those who did not develop a donor specific antibody had a class I allele only, while those who did develop a donor specific antibody were fairly evenly spread among class I, class II, or both. There is a significant relationship between allele class and development of a donor specific antibody ($p=0.01$).

```{r model}
df.blood$race2 <- factor(ifelse(df.blood$race_c=="Hispanic"|df.blood$race_c=="Asian"|df.blood$race_c=="No response","H/A/NR",df.blood$race_c))

mod <- glm(pra_bin ~ prior_bin + age + weight + prior_mech_nomiss +  no_of_prior_surgeries + gender_c + race2 + diag_c, data=df.blood, family="binomial",na.action = na.pass)

#step.aic <- step(mod)

finalmod <- glm(pra_bin ~ prior_bin + prior_mech_nomiss + race2 + diag_c,data=df.blood,family="binomial")


coefs <- summary(finalmod)$coefficients[,1]
coefs.se <- summary(finalmod)$coefficients[,2]
oddsrat <- exp(coefs)
or.se <- exp(confint(finalmod))
pvals <- summary(finalmod)$coefficients[,4]

mymat <- round(cbind(coefs,coefs.se,oddsrat,or.se,pvals),3)

tabmat <- matrix(0,nrow(mymat),3)
tabmat[,1] <- paste0(mymat[,1]," (",mymat[,2],")")
tabmat[,2] <- paste0(mymat[,3]," (",mymat[,4],",",mymat[,5],")")
tabmat[,3] <- mymat[,6]

rownames(tabmat) <- c("Intercept","Total blood products","Mech support","Race (AA)","Race (H/A/NR)","Diagnosis (Dilated CM)","Diagnosis (Failed SV)","Diagnosis (Restrictive CM)","Diagnosis (Other)")
kable(tabmat, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","Odds Ratio (95% CI)","p value"), caption = "Final Model") %>%
  kable_styling(latex_options = "hold_position")

library(pROC)
test_prob = predict(finalmod, type = "response")
test_roc = roc(df.blood$pra_bin ~ test_prob, plot = TRUE, print.auc = TRUE)

```

```{r modelparts}
mod2 <- glm(pra_bin ~ rbc_bin + plt_bin + ffp_bin + cryo_bin + prior_mech_nomiss + race2 + diag_c, data=df.blood, family="binomial",na.action = na.pass)

coefs <- summary(mod2)$coefficients[,1]
coefs.se <- summary(mod2)$coefficients[,2]
oddsrat <- exp(coefs)
or.se <- exp(confint(mod2))
pvals <- summary(mod2)$coefficients[,4]

mymat <- round(cbind(coefs,coefs.se,oddsrat,or.se,pvals),3)

tabmat <- matrix(0,nrow(mymat),3)
tabmat[,1] <- paste0(mymat[,1]," (",mymat[,2],")")
tabmat[,2] <- paste0(mymat[,3]," (",mymat[,4],",",mymat[,5],")")
tabmat[,3] <- mymat[,6]

rownames(tabmat) <- c("Intercept","Red blood cells","Platelets","Fresh frozen plasma","Cryoprecipitate","Mech support","Race (AA)","Race (H/A/NR)","Diagnosis (Dilated CM)","Diagnosis (Failed SV)","Diagnosis (Restrictive CM)","Diagnosis (Other)")
kable(tabmat, "latex", booktabs = T, align = "l", col.names = c("Estimate (SE)","Odds Ratio (95% CI)","p value"), caption = "Model with Individual Blood Components") %>%
  kable_styling(latex_options = "hold_position")

test_prob2 = predict(mod2, type = "response")
test_roc2 = roc(df.blood$pra_bin ~ test_prob2, plot = TRUE, print.auc = TRUE)



```



